A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-03-25 paper.tex
Hopefully cleared up some confusing wording in sect 2

The algorithm consists of two competing solvers, one for the system and one for
the environment, that solve abstractions of the safety game. Each solver builds
a candidate strategy for its corresponding player consisting of assignments to
$2^C$ and $2^U$ respectively.

@@

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
an example abstract game tree for the controller (abstract game trees for the
environment are similar).  In the abstract game, the controller is required to
pick actions from the tree, starting from the root node.  After reaching a
leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
restricts the initial controller action to the set $\{a, d\}$.  After choosing
action $a$, the controller is required to play action $b$ in the next round.
It then reaches a leaf of the tree and continues playing unrestricted.  The
root of the tree is annotated by the initial state $s$ of the abstract game and
the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
$n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a game
bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root node
of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs
$\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the edge
connecting $n$ and $n'$.

@@

The algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
argument.  In addition, it takes a player $p$ (controller or environment),
state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
invocation of the algorithm takes the initial state $I$, bound $\kappa$ and an
empty abstract game tree $\emptyset$.  Initially the solver is playing on
behalf of the environment since that player takes the first move in every game
round.  The empty game tree does not constrain opponent moves, hence solving
such an abstraction is equivalent to solving the original concrete game.

The algorithm is organised as a counterexample-guided abstraction refinement
(CEGAR) loop.  The first step of the algorithm uses the \textsc{findCandidate}
function, described below, to come up with a candidate partial strategy for
$T$. If it fails to find a strategy, this means that no winning partial
strategy exists for $T$.  If, on the other hand, a candidate partial strategy
is found, we need to verify if it is indeed winning for $T$.

@@

We extend the bounded synthesis algorithm to learn states losing for one of the
players from failed attempts to find candidate strategies.  The learning
procedure kicks in whenever \textsc{findCandidate} cannot find a candidate
strategy for an abstract game tree $T$ with initial state $s$ and bound $k$.
For the system, this means the environment partial strategy represented by $T$
will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
bound is irrelevant).  For the environment, the system can avoid $E$ for $k$
rounds no matter the assignments to $U$ variables. Hence, we have established
that $s$ is a losing state either always (for the system) or for bound $k$ (for
the environment. We can learn additional losing states from the tree via
interpolation.  This is achieved in lines~18--20 in
Algorithm~\ref{alg:bounded}, enabled in the unbounded version of the algorithm,
which invoke \textsc{learn} or \textsc{\textoverline{learn}} to learn
controller or environment losing states respectively
(Algorithm~\ref{alg:learn}).

>>>>

The algorithm consists of two competing solvers, one for the controller and one for
the environment, that solve abstractions of the safety game. Each solver builds
a candidate strategy for its corresponding player consisting of assignments to
$2^C$ and $2^U$ respectively.

@@

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
an example abstract game tree restricting the controller (abstract game trees
restricting the environment are similar).  In the abstract game, the controller
is required to pick actions from the tree, starting from the root node.  After
reaching a leaf, it continues playing unrestricted.  The tree in
Figure~\ref{fig:agt} restricts the initial controller action to the set $\{a,
d\}$.  After choosing action $a$, the controller is required to play action $b$
in the next round.  It then reaches a leaf of the tree and continues playing
unrestricted.  The root of the tree is annotated by the initial state $s$ of
the abstract game and the bound $k$ on the number of rounds.  We denote
$\textsc{nodes}(T)$ the set of all nodes of a tree $T$, $\textsc{leaves}(T)$
the subset of leaf nodes.  For edge $e$, $\textsc{action}(e)$ is the action
that labels the edge, and for node $n$, $\textsc{height}(k, n)$ is the distance
from n to the last round of a game bounded to $k$ rounds.  $\textsc{height}(k,
T)$ is the height of the root node of the tree.  For node $n$ of the tree,
$\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$ where $n'$ is a
child node of $n$ and $e$ is the edge connecting $n$ and $n'$.

@@

The algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
argument.  In addition, it takes a player $p$ (controller or environment),
state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
partial strategy for $p$, if one exists.  The initial invocation of the
algorithm takes the initial state $I$, bound $\kappa$ and an empty abstract
game tree $\emptyset$.  Initially the solver is playing on behalf of the
environment since that player takes the first move in every game round.  The
empty game tree does not constrain opponent moves, hence solving such an
abstraction is equivalent to solving the original concrete game.

The algorithm is organised as a counterexample-guided abstraction refinement
(CEGAR) loop.  The first step of the algorithm uses the \textsc{findCandidate}
function, described below, to come up with a candidate partial strategy that is
winning when the opponent is restricted to $T$.  If it fails to find a
strategy, this means that no winning strategy exists against the opponent
playing its partial strategy represented by $T$.  If, on the other hand, a
candidate partial strategy is found, we need to verify if it is indeed winning
for the abstract game $T$.

@@

We extend the bounded synthesis algorithm to learn states losing for one of the
players from failed attempts to find candidate strategies.  The learning
procedure kicks in whenever \textsc{findCandidate} cannot find a candidate
strategy for an abstract game tree $T$ with initial state $s$ and bound $k$.
For the controller, this means the environment partial strategy represented by $T$
will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
bound is irrelevant).  For the environment, the controller can avoid $E$ for $k$
rounds no matter the assignments to $U$ variables. Hence, we have established
that $s$ is a losing state either always (for the controller) or for bound $k$ (for
the environment. We can learn additional losing states from the tree via
interpolation.  This is achieved in lines~18--20 in
Algorithm~\ref{alg:bounded}, enabled in the unbounded version of the algorithm,
which invoke \textsc{learn} or \textsc{\textoverline{learn}} to learn
controller or environment losing states respectively
(Algorithm~\ref{alg:learn}).
